Distributed Adaptive Nrom Estimation For Blind System\\Identification in Wireless Sensor Networks

### 2022-10-21 paper.tex
Included Toon's feedback

% Title.
% ------
\title{Distributed averaging-based approach to BSI with ADMM}
%
% Single address.
% ---------------
\name{M. Blochberger\(^1\), R. Ali\(^1\), F. Elvander\(^2\), T. van Waterschoot\(^1\),tbd\thanks{This research work was carried out at the ESAT Laboratory of KU Leuven, in the frame of the SOUNDS European Training Network. This project has received funding from the European Union's Horizon 2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement No.\,956369. This research received funding in part from the Research Foundation - Flanders (FWO) grant 12ZD622N as well as from the European Union's Horizon 2020 research and innovation program / ERC Consolidator Grant: SONORA (No.\,773268). This paper reflects only the authors' views and the Union is not liable for any use that may be made of the contained information. Source code available at \textcolor{red}{ADD URL}}}
\address{\(^1\)KU Leuven, Dept. of Electrical Engineering (ESAT), STADIUS, 3001 Leuven, Belgium\\\(^2\)Aalto University, Dept. of Signal Processing and Acoustics, 02150 Espoo, Finland\\tbd}

\begin{document}
\ninept
%
\maketitle
%
\begin{abstract}
    Distributed signal-processing algorithms in (wireless) sensor networks often aim to decentralize processing tasks to reduce communication cost and computational complexity or avoid reliance on a single device for processing.
    In this contribution, we extend a distributed adaptive algorithm for blind system identification that relies on the knowledge of a stacked network-wide solution vector at each node, the computation of which requires either broadcasting or relaying of node-specific values to all other nodes.
    The extended algorithm employs a distributed-averaging-based estimation scheme to estimate the network-wide norm value by only using the information provided by neighboring sensor nodes.
    We introduce an adaptive mixing factor between instantaneous and recursion values for adaptivity in a time-varying system.
    Simulation results show that the extension provides estimation results close to the optimal fully-connected-network or broadcasting case while reducing inter-node transmission significantly.
\end{abstract}
%
\begin{keywords}
multi-channel signal processing, distributed signal processing, wireless sensor networks, blind system identification, distributed averaging
\end{keywords}
%
\section{INTRODUCTION}
\label{sec:intro}

Distributed algorithms have been an active area of research for quite some time, with numerous control, optimization, and signal processing applications.
With the ever-growing number of smart multimedia devices in today's surroundings providing ubiquitous processing and communication capabilities, distributed audio and speech signal processing have also found their way into the spotlight.
Algorithms for distributed signal estimation \cite{5483092}, noise control and echo cancellation \cite{9670697}, as well as beamforming \cite{6663655,6329934,MARKOVICHGOLAN20154} to name a few, have been proposed.
Another important task in audio and communication applications, which however has not recived as much attention, is multi-channel system identification, i.e. estimating channel responses in time or frequency domain.
Distributed single-input-multiple-output (SIMO) blind system identification (BSI) had contributions such as the adaptive cross-relation-based (CR) \cite{yuDistributedBlindSystem2014, liuDistributedBlindIdentification2016}.
Furthermore, we recently introduced an adaptive CR-based algorithm \cite{blochbergerDBSI} using the alternating direction method of multipliers (ADMM) \cite{boydDistributedOptimizationStatistical2011}.

@@

In this paper, we extend \cite{blochbergerDBSI} with a distributed averaging-based \cite{xiaoFastLinearIterations2004} estimation scheme for the global variable.
This approach allows us to compute the global variable without the need of a fully connected network or broadcasting.
We also introduce a mixing factor to include instantaneous values into the averaging recursion, which then allows us (i) to reduce the number of secondary iterations significantly (down to 1) and (ii) track time-varying systems.
Simulation results demonstrate how the extension leads to BSI performance close to an optimal case where all information is available without the need of a fully connected network or a large number of estimation iterations.

\section{DISTRIBUTED ADAPTIVE BSI IN SENSOR NETWORKS WITH ONLINE-ADMM}
\label{sec:dbsi}
In this section, we briefly outline the adaptive SIMO BSI algorithm proposed in [8], setting the scene for the distributed averaging scheme proposed herein.
% In this section, we will introduce certain parts of an adaptive SIMO BSI algorithm based on Online-ADMM \cite{blochbergerDBSI}, which are necessary for the reader to understand its extension of it.
We refer the reader to the publication cited above for the full derivation of the algorithm.

\subsection[]{Sensor network}
We assume a set \(\Mset \triangleq \{1,\ldots,M\}\) sensor nodes indices and a set of edge tuples \(\mathcal{E}\), which connect the nodes forming a sensor network.
Each edge is an unordered pair of node indices \(\{i,j\} \in \mathcal{E}\), which represents the communication link between them.
We allow \(\{i,i\} \in \mathcal{E}\), in order to simplify notation later, however, this does not represent a link but rather indicates the obvious fact that node \(i\) has information of itself.
Furthermore, let \(\Nset_i = \{j|\{i,j\} \in \mathcal{E}\}\) denote the neighborhood of node \(i\),
% From this follows the neighborhood of a node \(i\), i.e., the other nodes it is directly connected to, defined as \(\Nset_i = \{j|\{i,j\} \in \mathcal{E}\}\).
% It is to note, that we define the set with \(i \in \Nset_i\) for ease of notation later.
% This does however not represent an actual inter-node link.
We can define the symmetric adjacency matrix \(\mtxb{C}\) with elements \(C_{ij} = 1\) if \(\{i,j\} \in \mathcal{E}\) and 0 otherwise.
It may be noted that in \cite{blochbergerDBSI}, the pairs \(\{i,j\} \in \mathcal{E}\) are ordered, which yields a directed graph and leads to two sets of neighborhoods (``transmit'' and ``receive'') and therefore a non-symmetric adjacency matrix.
For the sake of brevity, we limit the explanations in this paper to the undirected case.
The case of directed graphs can be extended directly from the presented results.

\subsection[]{Signal model}
We consider a SIMO system with
\begin{align}
    % \textstyle
    \mtxb{s}(n) &= [s(n),\,\ldots,\,s(n-2L+2)]^{\T},\\
    \x_i(n) &= [x_i(n),\,\ldots,\,x_i(n-L+1)]^{\T}, \quad i \in \Mset,
\end{align}
the input signal frame and \(M\) output signal frames, respectively.
Each output \(\x_i(n)\) is the convolution of \(\mtxb{s}(n)\) with the respective channel impulse response \(\h_i\) and an additive noise term \(\mtxb{v}_i(n)\), assumed to be zero-mean and uncorrelated with \(\mtxb{s}(n)\).
The signal model is \(\x_i(n) = \mtxb{H}_i \mtxb{s}(n) + \mtxb{v}_i(n),\)
with \(\mtxb{H}_i\), the \(L \times (2L-1)\) linear convolution matrix of the \(i\)th channel using the elements of \(\h_i\) of length \(L\).
For the purpose of this paper, the length \(L\) of the impulse responses is assumed to be known.

\subsection[]{Distributed BSI with Online-ADMM}
In BSI, the coss-relation problem formulation only uses output signals \(\x_i(n)\), exploiting relative information between them, to identify the system, i.e., the acoustic or communication channels \(\h = [\h_1^\T,\,...,\,\h_M^\T]^\T\).
The solution to this problem is found by the minimization problem \cite{langtongBlindIdentificationEqualization1994,huangAdaptiveMultichannelLeast2002,huangClassFrequencydomainAdaptive2003,blochbergerDBSI}:
        \hat{\h} = \arg \min_{\h} \quad &\h^\herm \R \h \\
        \text{s.t. } \quad &\|\h\| = 1.
    \end{aligned}\label{eq:frequency_domain:min_prob}
where \(\hat{\h} = [\hat{\h}_1^\T,\,...,\,\hat{\h}_M^\T]^\T\) is the vector of estimated channel responses.
In \cite{blochbergerDBSI}, this problem is solved by a distributed adaptive algorithm that is based on separating the problem \eqref{eq:frequency_domain:min_prob} into node-wise subproblems,
        \underset{\w,\,\h}{\min} \quad &\sum_{i \in \Mset} \w_i^\herm \aRhof_i \w_i\\
        \text{s.t.} \quad &(\w_i)_j = \h_{\mathcal{G}(i,j)}\quad i \in \Mset,\,j \in \Nset_i\\
        &\|\h\| = 1.
    \end{aligned}\label{eq:general_consensus_admm:min_prob}
Here, each subproblem \(\w_i^\herm \aRhof_i \w_i\), with the node-local estimates \(\w_i\) and data matrix \(\aRhof_i\).
Both are analogous to \(\h_i\) and \(\R\) in \eqref{eq:frequency_domain:min_prob}, however, only using a subset channels, provided by neighboring nodes \(\Nset_i \subset \Mset\).
The first constraint enforces a consensus of local estimates, and \(\mathcal{G}(i,j)\) is a parameter mapping function equating local estimates of nodes connected by edge \((i,j) \in \mathcal{E}\).
General-form consensus alternating direction method of multipliers (ADMM) \cite{boydDistributedOptimizationStatistical2011} is applied in an adaptive updating scheme (Online-ADMM) \cite{wangOnlineAlternatingDirection2013,hosseiniOnlineDistributedADMM2014}.
ADMM leads to update steps for local, consensus, and dual variables.
We refer the reader to \cite{blochbergerDBSI} for details.
The algorithm's consensus variable is the system's estimate, \(\hat{\h}\).
Each node \(i\) computes the respective subvector \(\hat{\h}_i\) locally, which is the computation step of interest for this extension.
% Separating the cost function and applying general-form consensus ADMM as an adaptive algorithm leads to the update steps
% \begin{align}
%     \w_i^{(n+1)} &= \underset{\w_i}{\operatorname{argmin}} \, \mathcal{L}_{\rho} (\w,\h^{(n)},\uu^{(n)})\label{eq:general_consensus_admm:local}\\
%     \h^{(n+1)} &= \underset{\h, \|\h\| = 1}{\operatorname{argmin}}\, \mathcal{L}_{\rho} (\w^{(n+1)},\h,\uu^{(n)})\label{eq:general_consensus_admm:global}\\
%     \uu_i^{(n+1)} &= \uu_i^{(n)} + \rho \left( \w_i^{(n+1)} - \tilde{\h}_i^{(n+1)} \right),\label{eq:general_consensus_admm:dual}
% \end{align}
% where \(\w_i\) represents channels estimates of lower dimensional subproblems, solved locally by node \(i\), \(\h = [\hat{\h}_1^\T,\,...,\,\hat{\h}_M^\T]^\T\) the consensus variable (estimate of the system), and \(\uu_i\) are the dual variables.
% \begin{todo}
%     Is this enough? I fear i have to use a full page introducing all the local, consensus, dual variables, parameter mapping and the lagrangian etc. only to get to \eqref{eq:online_admm:consensus_update} and say that the denominator is what we are looking at...
% \end{todo}
It is defined as
    \hat{\h_i}^{(n+1)} = \frac{\bar{\h}_i^{(n+1)}}{\sqrt{\sum_{j \in \Mset} \|\bar{\h}_j^{(n+1)}\|^2}}\label{eq:online_admm:consensus_update}
for each node \(i \in \Mset\) with the local unnormalized consensus \(\bar{\h}_i^{(n+1)}\), a combination of neigborhood averages of local estimates and dual variables, see \cite{blochbergerDBSI}.
The computation of the denominator of \eqref{eq:online_admm:consensus_update} requires  \(\|\bar{\h}_j^{(n+1)}\|\) from all nodes \(j \in \Mset\), which however, is not possible without a fully-connected network or broadcasting.
In \cite{blochbergerDBSI}, we assume that partial squared norms are relayed throughout the network until all nodes have the information necessary.

@@

Subsequently, we introduce an extension to the algorithm described in \cite{blochbergerDBSI} using a fastest distributed linear averaging (FDLA) approach \cite{xiaoFastLinearIterations2004} to avoid the need of network wide data transmission.
The approach described in \cite{yuDistributedBlindSystem2014,liuDistributedBlindIdentification2016} employs a related iterative averaging scheme within a distributed gradient descent algorithm for BSI in order to estimate a global variable, however, necessitating many iterations per frame.
So we further introduce an adaptive mixing factor to include instantaneous values in the recursion to separate avaraging iterations onto time frames of the adaptive algorithm.

@@

\subsection[]{Distributed averaging}
% The computation of the denominator of \eqref{eq:online_admm:consensus_update} requires knowledge of all \(\|\bar{\h}_i\|^2\) at each node \(i \in \Mset\).
% To avoid the large number of transmissions, which would be necessary to facilitate this, we will approximate the squared norm \(\sum_{i \in \Mset} \|\bar{\h}_i\|^2\) at each node \(i\) only using information shared within its neighborhood \(\Nset_i\).
First, assuming we have some unnormalized consensus variables \(\bar{\h}_i\) for \(i \in \Mset\), we look to distributively compute a sum of their squared norms, \(\sum_{i \in \Mset} \|\bar{\h}_i\|^2\), for \eqref{eq:online_admm:consensus_update} at each node \(i\).

@@

They have the form of

@@

In this application, the elements of \(\bm{\phi}(0)\) are the initial node-wise squared norm values \(\|\bar{\h}_i\|^2\), of which we want to compute the sum of.
This leads to \(\lim_{k \to \infty} M \phi_i(k) = \sum_{i \in \Mset} \|\bar{\h}_i\|^2\) for all \(i \in \Mset\).

@@

It is to note that this has to be done only once for any fixed network \(\Mset\) and \(\mathcal{E}\).
% \begin{todo}
%     This section reads awkward
% \end{todo}
% In usual applications of the iteration \eqref{eq:adaptivenormest:distlincomb}, it is applied \(K \in \mathbb{N}\) times until convergence of the averages is reached.

@@

Given that the BSI algorithm being extended \cite{blochbergerDBSI} is adaptive and the unnormalized consensus estimates \(\bar{\h}_i^{(n)}\) are therefore time-varying, computing a distributed sum of their norms has to be adapted.

@@

For this iterative approach to converge, \(K\) has to be sufficiently large (e.g., \(K=50\) in \cite{yuDistributedBlindSystem2014,liuDistributedBlindIdentification2016}).

@@

Therefore, we first propose to split iterations over time frames, i.e., each \(n\), we set the initial values \(\phi_i^{(n)}(0) = \phi_i^{(n-1)}(K)\), that is, the result of the last iteration of the previous frame.
This means \(K\) can be chosen small, given that after sufficient frames \(n\), convergence will be reached.
This rudimentary first modification, however, does not factor in the adaptive approach to channel identification.
That is, when letting \(\phi_i^{(n)}(0) = \|\bar{\h}_i^{(0)}\|\), with \(\bar{\h}_i^{(0)}\) being estimates at the inital frame \(n=0\), the approach converges to the sum of these inital - bad - estimates.
Therefore we factor in the instantaneous approximation, the weighted in-neighborhood sum \(\eta_i^{(n)} = \sum_{j \in \Nset_i} W_{ij} \|\bar{\h}_i^{(n)}\|^2\).
This leads to

@@

    \hat{\phi}_i^{(n+1)} = \gamma_i \eta_i^{(n)} + (1-\gamma_i) \phi_i^{(n)}(K),

@@

% The combination of instantaneous values and distributed averaging estimate should lead to lower steady-state error when \(\gamma_i\) is chosen right, while with adaptive \(\gamma_i\), convergence speed should increase as well.
Refer to \autoref{alg:davg_norm_est} for the extended algorithm introduced in this section.

% \section{Transmission cost}

\section{Evaluation}
\label{sec:simulations}
\subsection[]{Communication cost}
\label{sec:transcost}
% \renewcommand{\arraystretch}{1.2}
% \begin{table}[t]
%     \centering
%     \begin{tabular}{ |l|l| } 
%         \hline
%         & Cost \\
%         \hline\hline
%         (1) Fully connected & \(\mathcal{O}(M)\) \\
%         \hline
%         (2) Broadcast & \(\mathcal{O}(M)\) \\ 
%         \hline
%         (2) Neighborhood & \(\mathcal{O}(\bar{N}_i K)\) \\ 
%         \hline
%     \end{tabular}
%     \caption[]{Comparison of communication cost.}
%     \label{tab:transcost:table}
% \end{table}
% \renewcommand{\arraystretch}{1.0}
To describe the communication cost within the network, we count the number of variables transmitted per time frame \(n\).
The most useful measure in the case of this work is to analyze the cost per node.
% For this we define an average neighborhood size \(\bar{N}_i = \frac{1}{M} \sum_{i \in \Mset} N_i\) where \(N_i = | \Nset_i |\) is the cardinality/size of the neigborhood set for node \(i\).
We compare the following communication schemes within the context of the algorithm \cite{blochbergerDBSI} this paper proposes the extension to:
\begin{itemize}
    \itemsep-0.2em
    \item[(1)] A fully-connected network, where node \(i\) communicates \(\|\bar{\h}_i\|^2\) to all other nodes \(\{j \in \Nset_i = \Mset\}\) directly, i.e., the neighborhood is the full network.
    % \item[(2)] The node \(i\) broadcasts \(\|\bar{\h}_i\|\), i.e., it transmits to all other nodes \(\{j \in \Mset \setminus i\}\) without direct links.
    \item[(2)] The node \(i\) communicates \(\|\bar{\h}_i\|^2\) only to neighboring nodes \(\{j \in \Nset_i \subset \Mset\}\) and applies \(K\) distributed averaging iterations.
\end{itemize}
\autoref{tab:transcost:table} compares the number of transmit and receive operations a node \(i\) has to apply in order to compute \eqref{eq:online_admm:consensus_update}.
Further, it shows the complexity of the problem a node \(i\) has to solve according to \cite{blochbergerDBSI} given the two types of network topology, assuming it uses all information it receives.
The advantage of the fully distributed algorithm becomes clear when \(M>>N_i\), as the additional communication and local complexity depends on neighborhood size \(N_i\) and not on network size \(M\).
% In this comparison, we only consider communication necessary for the computation of \eqref{eq:online_admm:consensus_update} as the rest of the algorithm is the same for our options.
% For (1), the cost per node increases linearly with network size \(M\), whereas for communication scheme (2), the cost is linearly dependent on neighborhood size \(N_i\) and iteration count \(K\) only and therefore independent of network size.
% The problem to be solved at each node \(i \in \Mset\) has complexity \(\mathcal{O}(M^2 L^2)\) for (1), with channel response length \(L\) (which means there is no reduction compared to a centralized approach).
% For (2), local complexity is \(\mathcal{O}(N_i^2 L^2)\), where \(N_i\) is the neighborhood size.
\vspace*{-0.6em}
\renewcommand{\arraystretch}{1.2}
\begin{table}[h]
    \begin{tabular}{ |l|l|l|l| }
        \hline
        & Trans. Ops. & Rec. Ops. & Local Compl. \\
        \hline\hline

>>>>

% Title.
% ------
\title{Distributed Adaptive Nrom Estimation For Blind System\\Identification in Wireless Sensor Networks}
%
% Single address.
% ---------------
\name{M. Blochberger\(^1\), R. Ali\(^1\), F. Elvander\(^2\), tbd, T. van Waterschoot\(^1\)\thanks{This research work was carried out at the ESAT Laboratory of KU Leuven, in the frame of the SOUNDS European Training Network. This project has received funding from the European Union's Horizon 2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement No.\,956369. This research received funding in part from the Research Foundation - Flanders (FWO) grant 12ZD622N as well as from the European Union's Horizon 2020 research and innovation program / ERC Consolidator Grant: SONORA (No.\,773268). This paper reflects only the authors' views and the Union is not liable for any use that may be made of the contained information. Source code available at \textcolor{red}{ADD URL}}}
\address{\(^1\)KU Leuven, Dept. of Electrical Engineering (ESAT), STADIUS, 3001 Leuven, Belgium\\\(^2\)Aalto University, Dept. of Signal Processing and Acoustics, 02150 Espoo, Finland\\tbd}

\begin{document}
\ninept
%
\maketitle
%
\begin{abstract}
    Distributed signal-processing algorithms in (wireless) sensor networks often aim to decentralize processing tasks to reduce communication cost and computational complexity or avoid reliance on a single device (i.e., fusion center) for processing.
    In this contribution, we extend a distributed adaptive algorithm for blind system identification that relies on the estimation of a stacked network-wide consensus vector at each node, the computation of which requires either broadcasting or relaying of node-specific values (i.e., local vector norms) to all other nodes.
    The extended algorithm employs a distributed-averaging-based estimation scheme to estimate the network-wide consensus norm value by only using the local vector norm provided by neighboring sensor nodes.
    We introduce an adaptive mixing factor between instantaneous and recursive estimates of these norms for adaptivity in a time-varying system.
    Simulation results show that the extension provides estimation results close to the optimal fully-connected-network or broadcasting case while reducing inter-node transmission significantly.
\end{abstract}
%
\begin{keywords}
multi-channel signal processing, distributed signal processing, wireless sensor networks, blind system identification, distributed averaging
\end{keywords}
%
\section{INTRODUCTION}
\label{sec:intro}

Distributed algorithms have been an active area of research for quite some time, with numerous control, optimization, and signal processing applications.
With the ever-growing number of smart multimedia devices in today's surroundings providing ubiquitous processing and communication capabilities, distributed audio and speech signal processing have also found their way into the spotlight.
Algorithms for distributed signal estimation \cite{5483092}, noise control and echo cancellation \cite{9670697}, as well as beamforming \cite{6663655,6329934,MARKOVICHGOLAN20154,6309434} to name a few, have been proposed.
Another important task in audio and communication applications, which however has not recived as much attention, is multi-channel system identification, i.e. estimating channel responses in the time or frequency domain.
Distributed single-input-multiple-output (SIMO) blind system identification (BSI) has been addressed with adaptive cross-relation-based (CR) algorithms \cite{yuDistributedBlindSystem2014, liuDistributedBlindIdentification2016}.
In this context, we recently introduced an adaptive CR-based algorithm \cite{blochbergerDBSI} using the alternating direction method of multipliers (ADMM) \cite{boydDistributedOptimizationStatistical2011}.

@@

In this paper, we extend the algorithm of \cite{blochbergerDBSI} with a distributed averaging-based \cite{xiaoFastLinearIterations2004} estimation scheme for the global variable.
This approach allows us, similar to further approaches like \cite{6334305,9914798}, to compute the global variable without the need of a fully connected network or broadcasting.
We also introduce a mixing factor to include instantaneous values into the averaging recursion, which then allows us (i) to reduce the number of secondary iterations significantly (down to 1) and (ii) track time-varying systems.
Simulation results demonstrate how the extension leads to BSI performance close to an optimal case where all information is available without the need of a fully connected network or a large number of estimation iterations.

\section{DISTRIBUTED ADAPTIVE BSI IN SENSOR NETWORKS WITH ONLINE-ADMM}
\label{sec:dbsi}
In this section, we briefly outline the adaptive SIMO BSI algorithm proposed in \cite{blochbergerDBSI}, setting the scene for the distributed averaging scheme proposed herein.
We refer the reader to the publication cited above for the full derivation of the algorithm.

\subsection[]{Sensor network}
We assume a set of nodes with one sensor each, that is, each node provides one signal.
Let this set of node be \(\Mset \triangleq \{1,\ldots,M\}\) and \(\mathcal{E}\) be a set of edge tuples, which connect the nodes forming a sensor network.
Each edge is an unordered pair of node indices \(\{i,j\} \in \mathcal{E}\), which represents the communication link between them.
We allow \(\{i,i\} \in \mathcal{E}\), in order to simplify notation later, however, this does not represent a link but rather indicates the obvious fact that node \(i\) has information of itself.
Furthermore, let \(\Nset_i = \{j|\{i,j\} \in \mathcal{E}\}\) denote the neighborhood of node \(i\).
We can define the symmetric adjacency matrix \(\mtxb{C}\) with elements \(C_{ij} = 1\) if \(\{i,j\} \in \mathcal{E}\) and 0 otherwise.
It may be noted that in \cite{blochbergerDBSI}, the pairs \(\{i,j\} \in \mathcal{E}\) are ordered, which yields a directed graph and leads to two sets of neighborhoods (``transmit'' and ``receive'') and therefore a non-symmetric adjacency matrix.
For the sake of brevity, we limit the explanations in this paper to the undirected case.
The case of directed graphs can be extended directly from the presented results.

\subsection[]{Signal model}
We consider a SIMO system with
\begin{align}
    % \textstyle
    \mtxb{s}(n) &= [s(n),\,\ldots,\,s(n-2L+2)]^{\T},\\
    \x_i(n) &= [x_i(n),\,\ldots,\,x_i(n-L+1)]^{\T}, \quad i \in \Mset,
\end{align}
the \(2L \times 1\) input signal frame and \(M\) \(L \times 1\)  output signal frames, respectively.
Each output \(\x_i(n)\) is the convolution of \(\mtxb{s}(n)\) with the respective channel impulse response \(\h_i\) and an additive noise term \(\mtxb{v}_i(n)\), assumed to be zero-mean and uncorrelated with \(\mtxb{s}(n)\).
The signal model is \(\x_i(n) = \mtxb{H}_i \mtxb{s}(n) + \mtxb{v}_i(n),\)
with \(\mtxb{H}_i\), the \(L \times (2L-1)\) linear convolution matrix of the \(i\)th channel using the elements of \(\h_i\) of length \(L\).
For the purpose of this paper, the length \(L\) of the impulse responses is assumed to be known.

\subsection[]{Distributed BSI with Online-ADMM}
In BSI, the cross-relation problem formulation only uses output signals \(\x_i(n)\), exploiting relative information between them, to identify the system, i.e., the acoustic or communication channels \(\h = [\h_1^\T,\,...,\,\h_M^\T]^\T\).
The solution to this problem is found by the minimization problem \cite{langtongBlindIdentificationEqualization1994,huangAdaptiveMultichannelLeast2002,huangClassFrequencydomainAdaptive2003,blochbergerDBSI}:
\begin{equation}
    \begin{aligned}
        \hat{\h} = \arg \min_{\h} \quad &\h^\herm \R \h \\
        \text{s.t. } \quad &\|\h\| = 1.
    \end{aligned}\label{eq:frequency_domain:min_prob}
\end{equation}
where \(\hat{\h} = [\hat{\h}_1^\T,\,...,\,\hat{\h}_M^\T]^\T\) is the vector of estimated channel responses and \(\R\) is the so-called cross-relation (CR) matrix.
In \cite{blochbergerDBSI}, this problem is solved by a distributed adaptive algorithm that is based on separating the problem \eqref{eq:frequency_domain:min_prob} into node-wise subproblems,
\begin{equation}
    \begin{aligned}
        \underset{\w,\,\h}{\min} \quad &\sum_{i \in \Mset} \w_i^\herm \aRhof_i \w_i\\
        \text{s.t.} \quad &(\w_i)_j = \h_{\mathcal{G}(i,j)}\quad i \in \Mset,\,j \in \Nset_i\\
        &\|\h\| = 1.
    \end{aligned}\label{eq:general_consensus_admm:min_prob}
\end{equation}
Here, each subproblem is represented by a cost function \(\w_i^\herm \aRhof_i \w_i\), with the node-local channel estimates \(\w_i\) and CR matrix \(\aRhof_i\).
Both are analogous to \(\h_i\) and \(\R\) in \eqref{eq:frequency_domain:min_prob}, however, only using a subset channels, corresponding to \(\Nset_i \subset \Mset\).
The first constraint enforces a consensus of local estimates, and \(\mathcal{G}(i,j)\) is a parameter mapping function equating local estimates of nodes connected by edge \((i,j) \in \mathcal{E}\).
The general-form consensus alternating direction method of multipliers (ADMM) \cite{boydDistributedOptimizationStatistical2011} is applied in an adaptive updating scheme (Online-ADMM) \cite{wangOnlineAlternatingDirection2013,hosseiniOnlineDistributedADMM2014}.
ADMM leads to update steps for local, consensus, and dual variables.
We refer the reader to \cite{blochbergerDBSI} for details.
The algorithm's consensus variable is the system's estimate, \(\hat{\h}\).
Each node \(i\) computes the respective subvector \(\hat{\h}_i\) locally, which is the computation step of interest for this extension.
It is defined as
\begin{equation}
    \hat{\h_i}^{(n)} = \frac{\bar{\h}_i^{(n)}}{\sqrt{\sum_{j \in \Mset} \|\bar{\h}_j^{(n)}\|^2}}\label{eq:online_admm:consensus_update}
\end{equation}
for each node \(i \in \Mset\) with the local unnormalized consensus \(\bar{\h}_i^{(n)}\), a combination of neigborhood averages of local estimates and dual variables, see \cite{blochbergerDBSI}.
The computation of the denominator of \eqref{eq:online_admm:consensus_update} requires  \(\|\bar{\h}_j^{(n)}\|\) from all nodes \(j \in \Mset\), which however, is not possible without a fully-connected network or broadcasting.
In \cite{blochbergerDBSI}, we assume that partial squared norms are relayed throughout the network until all nodes have the information necessary.

@@

Subsequently, we introduce an extension to the algorithm described in \cite{blochbergerDBSI} using a fastest distributed linear averaging (FDLA) approach \cite{xiaoFastLinearIterations2004} to avoid the need of network-wide data transmission.
The approach described in \cite{yuDistributedBlindSystem2014,liuDistributedBlindIdentification2016} employs a related iterative averaging scheme within a distributed gradient descent algorithm for BSI in order to estimate a global variable, however, necessitating many iterations per frame.
To alleviate this, we further introduce an adaptive mixing factor to include instantaneous values in the recursion to separate avaraging iterations onto time frames of the adaptive algorithm.

@@

\subsection[]{Distributed averaging}
First, assuming we have some unnormalized consensus variables \(\bar{\h}_i\) for \(i \in \Mset\), we look to distributively compute a sum of their squared norms, \(\sum_{j \in \Mset} \|\bar{\h}_j\|^2\), for \eqref{eq:online_admm:consensus_update} at each node \(i\).

@@

They generally have the form of

@@

In this application, the elements of \(\bm{\phi}(0)\) are the initial node-wise squared norm values \(\|\bar{\h}_i\|^2\), which we want to compute the sum of.
This leads to \(\lim_{k \to \infty} M \phi_i(k) = \sum_{j \in \Mset} \|\bar{\h}_j\|^2\) for all \(i \in \Mset\).

@@

It is to note that this has to be done only once for any fixed-topology network \(\Mset\) and \(\mathcal{E}\).

@@

\begin{figure}[t]
    \input{simulations/plots/icassp2023-static-time/icassp2023-static-time.pgf}
    \vspace*{-0.6cm}
    \caption[]{Median NPM of 30 Monte-Carlo runs over time for optimal case where all node information is available (fully connected network/broadcasting), fixed values of \(\gamma_i\) with \(K \in \{1,10\}\), adaptive \(\gamma_i\) with \(K=1\). A moving average filter was applied to result curves for better readability (compare transparent and opaque lines). Hatched area indicates frames used in computations for \autoref{fig:simulations:avgNPMgamma}.}
    \label{fig:simulations:NPMtime}
\end{figure}
Given that the BSI algorithm being extended \cite{blochbergerDBSI} is adaptive and the unnormalized consensus estimates \(\bar{\h}_i^{(n)}\) are therefore time-varying, computing a distributed sum of their norms has to be performed adaptively.

@@

For iterative approach to converge, \(K\) has to be sufficiently large (e.g., \(K=50\) in \cite{yuDistributedBlindSystem2014,liuDistributedBlindIdentification2016}).

@@

Therefore, we first propose to split iterations over time frames.
Each \(n\), we could set the initial values \(\phi_i^{(n)}(0)\) to the result of the \(K\)-th iteration \(\phi_i^{(n-1)}(K)\) of the previous frame \(n-1\).
With sufficiently large \(n\), convergence will be reached even when \(K\) is chosen small.
However, this rudimentary modification does not factor in the adaptive approach to channel identification.
That is, when letting \(\phi_i^{(0)}(0) = \|\bar{\h}_i^{(0)}\|\), with \(\bar{\h}_i^{(0)}\) being estimates at the inital frame \(n=0\), it converges to the sum of these initial - bad - estimates.
Therefore, we factor in the instantaneous approximation, the weighted in-neighborhood sum \(\eta_i^{(n)} = \sum_{j \in \Nset_i} W_{ij} \|\bar{\h}_i^{(n)}\|^2\), which will introduce new information gained by more accurate estimates of \(\bar{\h}_i^{(n)}\).
We add this information by linearly combining it with the \(K\)-th iterate of \(\phi_i^{(n)}(k)\) and set it as the initial value for the next frame:

@@

    \phi_i^{(n)}(0) = \gamma_i \eta_i^{(n)} + (1-\gamma_i) \phi_i^{(n-1)}(K),

@@

Refer to \autoref{alg:davg_norm_est} for the order of computations of the extended algorithm we introduced here.

\section{Evaluation}
\label{sec:simulations}
\begin{figure}[t]
    \centering
    \input{simulations/plots/icassp2023-static/icassp2023-static.pgf}
    \vspace*{-0.6cm}
    \caption[]{Mean of 500 post-convergence frames of median NPM of 30 Monte-Carlo runs for fixed \(\gamma_i\) on the interval \([0.0, 0.04]\) and \(K \in \{1,2,10\}\). Results for optimal algorithm, and adaptive \(\gamma_i\) included for comparison.}
    \label{fig:simulations:avgNPMgamma}
\end{figure}
\subsection[]{Communication cost}
\label{sec:transcost}
To describe the communication cost within the network, we count the number of variables transmitted per time frame \(n\).
The most useful measure in the case of this work is to analyze the cost per node.
We compare the following communication schemes within the context of the algorithm \cite{blochbergerDBSI} this paper proposes the extension to:
\begin{itemize}
    \itemsep-0.2em
    \item[(1)] A fully-connected network, where node \(i\) communicates \(\|\bar{\h}_i\|^2\) to all other nodes \(\{j \in \Nset_i = \Mset\}\) directly, i.e., the neighborhood is the full network.
    \item[(2)] The node \(i\) communicates \(\|\bar{\h}_i\|^2\) only to neighboring nodes \(\{j \in \Nset_i \subset \Mset\}\) and applies \(K\) distributed averaging iterations.
\end{itemize}
\autoref{tab:transcost:table} compares the number of transmit and receive operations a node \(i\) has to apply in order to compute \eqref{eq:online_admm:consensus_update}.
Further, it shows the complexity of the problem a node \(i\) has to solve according to \cite{blochbergerDBSI} given the two types of network topology, assuming it uses all information it receives.
The advantage of the fully distributed algorithm becomes clear when \(M \gg N_i\), as the additional communication and local complexity depends on neighborhood size \(N_i\) and not on network size \(M\).
\vspace*{-0.6em}
\renewcommand{\arraystretch}{1.2}
\begin{table}[h]
    \centering
    \begin{tabular}{ |l|l|l|l| }
        \hline
        & Trans. Ops. & Rec. Ops. & Local Compl. \\
        \hline\hline
        (1) Fully con. & \(M-1\) & \(M-1\) & \(\mathcal{O}(M^2 L^2)\) \\
        \hline
        (2) Neighborh. & \(N_i K\) & \(N_i K\) & \(\mathcal{O}(N_i^2 L^2)\)\\ 
        \hline
    \end{tabular}
    \caption[]{Communication cost per node and frame \(n\) for computation of \eqref{eq:online_admm:consensus_update}.}
    \label{tab:transcost:table}
\end{table}
\renewcommand{\arraystretch}{1.0}

\subsection[]{Simulations}
To evaluate the effectiveness of the proposed extension, simulations were run, where we define the error measure as the normalized projection misalignment
\begin{equation}
    \begin{aligned}
        \text{NPM}(n) &= 20\,\log_{10} \frac{\left\| \mtxb{e} \right\|}{\left\|\hat{\h}^{(n)}\right\|},\\
        \text{with}\quad \mtxb{e} &= \hat{\h}^{(n)} - \frac{\h^\T \hat{\h}^{(n)}}{\h^\T \h}\hat{\h}^{(n)},
    \end{aligned}
\end{equation}
commonly used in BSI to compare the estimated \(\hat{\h}^{(n)}\) and true \(\h\).
